Evaluating an epidemiologically motivated surrogate model of a multi-model ensemble

### 2023-01-09 paper/paper.Rmd
add additional paper changes]

To understand the behaviour of the Forecast Hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their forecast models and then use them to produce a weekly forecast that they then submit to the central hub team. 
Each hub targets a range of metrics, including test-positive reported cases, reported deaths, and hospitalisations; has a specific geographic focus, and asks for weekly forecasts (using MMWR epidemiological weeks i.e. Sunday to Saturday [@MMWRweek]) over a time horizon of a few weeks.
Observed data are available and updated daily, and most submitted forecasts use this dataset, along with potentially other sources of real-time information, to produce forecasts.
Here, we focus on reported cases and primarily on the European Forecast Hub but our observations hold, in our view, across COVID-19 Forecast Hubs and to a lesser degree targets.
We focus on reported cases as these represent the most common forecast target for COVID-19 forecast models [@Nixon2022-zv], they are often of the most direct interest due to being a leading indicator for other metrics such as hospitalisations [@Meakin2022-hw], and they are generally the most challenging to predict [@Sherratt2022-ue].
In general, 5 main classes of forecast models are submitted [@Bracher2022-ug; @Cramer2022-sp], statistical forecasting models such as ARIMA models, mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@Srivastava2020-qv; @Li2021-cj], semi-mechanistic approaches that blend both of these approaches [@Castro2021-lj; @Bosse2022-hm], agent-based simulation models [@Rakowski2010-jn; @Adamik2020-if], and human insight based forecast models that may also include elements of other methods [@Karlen2020-pt; @Bosse2022-hm].
Real-time evaluation has shown that each of these classes of models may perform well depending on the context and specific implementation of the forecast model [@Bosse2022-hm].

We extracted forecasts and data on notified weekly COVID-19 cases from the European forecasting hub [@Sherratt2022-ue; @EuroHub] from the 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth` and defined in the following section).
We included all locations covered by the European forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
Data on notified weekly cases was originally sourced from the Johns Hopkins University (JHU) curated data repository [@JHU].
We used the latest available observed data as of the 1st of September 2022 (commit `f6922c3e4bdcb055abcbba8e73472afacac4cf40` from [@EuroHubGitHub]).
Incidence was aggregated by epidemiological week (i.e. Sunday through Saturday).
As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
To account for this we followed the practice of the European forecasting hub project in excluding forecasts made using anomalous truth data in the week of the forecasts production and excluding forecasts for target weeks with anomalous data [@Sherratt2022-ue].
We defined anomalous data based on the implementation used by [@EuroHub] where a data point is considered anomalous if a future revision alters it by more than 5%.

@@

We have contributed a range of forecasts to COVID-19 forecast collaborations, generally focused on semi-mechanistic statistical methods and human insight-based forecasts.
Our forecast submissions have not systematically over- or under-performed other forecasts submitted to the forecasting hubs (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm; @Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue]).
The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
Based on our observations our forecasts have generally captured the current trend relatively well but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns. 
Our previous methodology also requires significant computational resources, running for an hour on a Azure D v5-series 16-core machine, when producing forecasts for the European forecasting hub [@AzurePricing]. 
This resource usage is likely beyond the capacity of many interested in having access to state-of-the-art short-term forecasts of infectious diseases.
In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
We found that our human insight-based forecasts outperformed our model-based forecasts on average. 
This was particularly the case when forecasting cases and at longer forecast horizons.
We hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@Bosse2022-hm].

@@

To download and manipulate forecasts from the European forecasting hub [@EuroHub] we use the `data.table` (`1.14.2`) [@data.table] and `gh` (`1.3.0`) [@gh] R packages.
We make use of further functionality from the `forecast.vocs` R package [@forecast.vocs] to prepare data for forecasting, visualise forecasts and summary measures, and summarise forecasts.
Forecast evaluation is implemented using the `scoringutils` R package (1.0.0) [@scoringutils], and the `scoringRules` R package (1.0.1) [@scoringRules].

@@

We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the European forecasting hub (nations of the European Union, the European Free Trade Agreement, and the United Kingdom, making `r locations` unique locations).
In total `r nrow(metadata)` forecasts were made across all locations, with `r forecast_dates` unique forecast dates and `r models` independent forecast models (including the European hub baseline model).
Of these models, `r length(global_models)` forecasted in at least 30 locations including our original submission  (referred to as `epiforecasts-EpiNow2` by the hub), and our surrogate model.
Of the remaining models submitted `r length(local_models)` were submitted in only one location.
Single-location models were clustered in a few locations, particularly in Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the European forecasting hub project [@Sherratt2022-ue]).
Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else.
`r nrow(partial_models)` models were submitted for between 3 and 30 locations and all these models varied the number of locations they submitted forecasts for over time, potentially indicating manual curation or models targeted at specific conditions.

@@

The poor calibration of the forecast ensembles produced by median Hub ensembles has been noted repeatedly [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue] but little progress has been made in understanding the causes or suggesting alternatives.
Progress in understanding which structural model features lead to better infectious disease forecasts has been limited.
The US Forecast Hub identified the top 5 performing models and noted the structural assumptions they made, but couldn't directly link assumptions with performance [@Cramer2022-sp].
They also did not extensively compare and contrast these conclusions to arrive at a set of desired forecast assumptions (as done in this study to motivate the surrogate model), or explore the performance of a forecasting model designed with these assumptions in mind.
Similarly, the Germany and Poland forecasting hubs were able to identify forecast models that performed comparably as well as their ensemble forecasts but did not derive structural assumptions that led to this out-performance or detail explicitly what the desirable performance characteristics would be, aside from optimising the weighted interval score.
All comparable Forecast Hub projects found that their ensemble was often the best choice, had desirable characteristics such as robustness - though this was rarely fully defined - and should be the output used by forecast consumers [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
In general, during the study period, all projects used the same unweighted median ensemble forecast of all submissions. 
The US [@Ray2022-zq], and European [@Sherratt2022-ue], forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensembling.
Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases in the case of the US forecasting hub and did not outperform in the case of the European forecasting hub.
No Forecast Hub has switched to these alternative ensemble designs for their operational forecast of reported cases, though the US hub has switched to a trained ensemble for death forecasts. This suggests that the hub teams do not think the evidence base is strong enough for trained ensembles to be used by forecast consumers for reported cases and hence the median of all submitted forecasts remains the community-suggested default ensemble option and a sensible target for our study.

>>>>

To understand the behaviour of the Forecast Hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their forecast models and then use them to produce a weekly forecast that they then submit to the central hub team. 
Each hub targets a range of metrics, including test-positive reported cases, reported deaths, and hospitalisations; has a specific geographic focus, and asks for weekly forecasts (using MMWR epidemiological weeks i.e. Sunday to Saturday [@MMWRweek]) over a time horizon of a few weeks.
Observed data are available and updated daily, and most submitted forecasts use this dataset, along with potentially other sources of real-time information, to produce forecasts.
Here, we focus on reported cases and primarily on the European Forecast Hub but our observations hold, in our view, across COVID-19 Forecast Hubs and to a lesser degree targets.
We focus on reported cases as these represent the most common forecast target for COVID-19 forecast models [@Nixon2022-zv], they are often of the most direct interest due to being a leading indicator for other metrics such as hospitalisations [@Meakin2022-hw], and they are generally the most challenging to predict as they do not have a clear leading indicator [@Sherratt2022-ue].
In general, 5 main classes of forecast models are submitted [@Bracher2022-ug; @Cramer2022-sp], statistical forecasting models such as ARIMA models, mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@Srivastava2020-qv; @Li2021-cj], semi-mechanistic approaches that blend both of these approaches [@Castro2021-lj; @Bosse2022-hm], agent-based simulation models [@Rakowski2010-jn; @Adamik2020-if], and human insight based forecast models that may also include elements of other methods [@Karlen2020-pt; @Bosse2022-hm].
Real-time evaluation has shown that each of these classes of models may perform well depending on the context and specific implementation of the forecast model [@Bosse2022-hm].

We extracted forecasts and data on notified weekly COVID-19 cases from the European Forecasting Hub [@Sherratt2022-ue; @EuroHub] from the 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth` and defined in the following section).
We included all locations covered by the European Forecasting Hub, which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
Data on notified weekly cases were originally sourced from the Johns Hopkins University (JHU) curated data repository [@JHU].
We used the latest available observed data as of the 1st of September 2022 (commit `f6922c3e4bdcb055abcbba8e73472afacac4cf40` from [@EuroHubGitHub]).
Incidence was aggregated by epidemiological week (i.e. Sunday through Saturday).
As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
To account for this we followed the practice of the European Forecasting Hub project in excluding forecasts made using anomalous truth data in the week of the forecasts production and excluding forecasts for target weeks with anomalous data [@Sherratt2022-ue].
We defined anomalous data based on the implementation used by [@EuroHub] where a data point is considered anomalous if a future revision alters it by more than 5%.

@@

We have contributed a range of forecasts to COVID-19 forecast collaborations, generally focused on semi-mechanistic statistical methods and human insight-based forecasts.
Our forecast submissions have not systematically over- or under-performed other forecasts submitted to the forecasting hubs (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm; @Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue]).
The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
Based on our observations our forecasts have generally captured the current trend relatively well but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns. 
Our previous methodology also requires significant computational resources, running for an hour on a Azure D v5-series 16-core machine, when producing forecasts for the European Forecasting Hub [@AzurePricing]. 
This resource usage is likely beyond the capacity of many interested in having access to state-of-the-art short-term forecasts of infectious diseases.
In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
We found that our human insight-based forecasts outperformed our model-based forecasts on average. 
This was particularly the case when forecasting cases and at longer forecast horizons.
We hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@Bosse2022-hm].

@@

To download and manipulate forecasts from the European Forecasting Hub [@EuroHub] we use the `data.table` (`1.14.2`) [@data.table] and `gh` (`1.3.0`) [@gh] R packages.
We make use of further functionality from the `forecast.vocs` R package [@forecast.vocs] to prepare data for forecasting, visualise forecasts and summary measures, and summarise forecasts.
Forecast evaluation is implemented using the `scoringutils` R package (1.0.0) [@scoringutils], and the `scoringRules` R package (1.0.1) [@scoringRules].

@@

We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the European Forecasting Hub (nations of the European Union, the European Free Trade Agreement, and the United Kingdom, making `r locations` unique locations).
In total `r nrow(metadata)` forecasts were made across all locations, with `r forecast_dates` unique forecast dates and `r models` independent forecast models (including the European hub baseline model).
Of these models, `r length(global_models)` forecasted in at least 30 locations including our original submission  (referred to as `epiforecasts-EpiNow2` by the hub), and our surrogate model.
Of the remaining models submitted `r length(local_models)` were submitted in only one location.
Single-location models were clustered in a few locations, particularly in Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the European Forecasting Hub project [@Sherratt2022-ue]).
Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else.
`r nrow(partial_models)` models were submitted for between 3 and 30 locations and all these models varied the number of locations they submitted forecasts for over time, potentially indicating manual curation or models targeted at specific conditions.

@@

The poor calibration of the forecast ensembles produced by median Hub ensembles has been noted repeatedly [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue] but little progress has been made in understanding the causes or suggesting alternatives.
Progress in understanding which structural model features lead to better infectious disease forecasts has been limited.
The US Forecast Hub identified the top 5 performing models and noted the structural assumptions they made, but couldn't directly link assumptions with performance [@Cramer2022-sp].
They also did not extensively compare and contrast these conclusions to arrive at a set of desired forecast assumptions (as done in this study to motivate the surrogate model), or explore the performance of a forecasting model designed with these assumptions in mind.
Similarly, the Germany and Poland forecasting hubs were able to identify forecast models that performed comparably as well as their ensemble forecasts but did not derive structural assumptions that led to this out-performance or detail explicitly what the desirable performance characteristics would be, aside from optimising the weighted interval score.
All comparable Forecast Hub projects found that their ensemble was often the best choice, had desirable characteristics such as robustness - though this was rarely fully defined - and should be the output used by forecast consumers [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
In general, during the study period, all projects used the same unweighted median ensemble forecast of all submissions. 
The US [@Ray2022-zq], and European [@Sherratt2022-ue], forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensembling.
Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases in the case of the US forecasting hub and did not outperform in the case of the European Forecasting Hub.
No Forecast Hub has switched to these alternative ensemble designs for their operational forecast of reported cases, though the US hub has switched to a trained ensemble for death forecasts. This suggests that the hub teams do not think the evidence base is strong enough for trained ensembles to be used by forecast consumers for reported cases and hence the median of all submitted forecasts remains the community-suggested default ensemble option and a sensible target for our study.
